Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-04-15 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@555 3fc257bb-8550-0410-81b3-a5c103da83f7

An ``independent'' set of vertices in a graph is one where no pair of
vertices in the set share an edge.  The set is ``maximal'' if no
additional vertex can be added to it \footnote{A maximal set is
``maximum'' if it is the largest maximal set.  Finding a graph's
maximum independent set is an NP-hard problem, which MapReduce is
unlikely to help with.}.  Finding a maximal independent set is useful
in several contexts, such as identifying large numbers of indepenent
starting vertices for graph traversals.  Luby's algorithm \cite{Luby}
is a well-known parallel method for finding a minimal set.  Figure
\ref{fig:luby} outlines a MapReduce version of Luby's algorithm.

@@

 \caption{MapReduce algorithm for finding amaximal independent set of
 graph vertices.}

@@

>>>>

An ``independent'' set of vertices in a graph is one where no pair of
vertices in the set share an edge.  The set is ``maximal'' if no
additional vertex can be added \footnote{A maximal set is ``maximum''
if it is the largest maximal set.  Finding a graph's maximum
independent set is an NP-hard problem, which MapReduce is unlikely to
help with.}.  Finding a maximal independent set (MIS) is useful in
several contexts, such as identifying large numbers of indepenent
starting vertices for graph traversals.  Luby's algorithm \cite{Luby}
is a well-known parallel method for finding a MIS.  Figure
\ref{fig:luby} outlines a MapReduce version of Luby's algorithm,
assuming an input graph of undirected edges $(V_i,V_j)$ where $V_i <
V_j$ for every edge, i.e. an upper-triangular R-MAT matrix.

\begin{figure}[htb]
 \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
  xxxx\=xxxxxxxxxxx\=xxxx\=xxxx\=xxxx\=xxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

Assign random values to each vertex \\
Clone: convert edge key/values directly to key/multivalues \\
Create empty MapReduce MRv for maximal independent set vertices \\
\\
while $N_{edges} > 0$: \\
\> 1 Reduce: \> Determine winner/loser vertex of each edge \\
       \> \> \> input Key = $E_{ij}$, Multivalue = NULLS or LOSER \\
       \> \> \> if any value is LOSER, emit nothing \\
       \> \> \> $V_w$ = WINNER vertex, $V_l$ = LOSER vertex \\
       \> \> \> output Key = $V_w$, Value = $V_l$ WINNER \\
       \> \> \> output Key = $V_l$, Value = $V_w$ LOSER \\
\> 1 Collate: \> Vertex as key \\
\> 2 Reduce: \> Find WINNER vertices \\
       \> \> \> input Key = $V_i$, MultiValue = $V V V V ...$ \\
       \> \> \> if all $V$ are WINNERs: \\
      \> \> \> \> output to MRv, Key = $V_i$, Value = NULL \\
      \> \> \> \> for each $V$ in Multivalue: \\
     \> \> \> \> \> output Key = $V_j$, Value = $Vi$ LOSER \\
       \> \> \> else: \\
      \> \> \> \> for each $V$ in Multivalue: \\
     \> \> \> \> \> output Key = $V_j$, Value = $V_i$ \\
\> 2 Collate: \> Vertex as key \\
\> 3 Reduce: \> Find LOSER vertices, delete their edges \\
       \> \> \> input Key = $V_i$, MultiValue = $V V V V ...$ \\
       \> \> \> if any $V$ is LOSER: \\
      \> \> \> \> for each $V$ in Multivalue: \\
     \> \> \> \> \> output Key = $E_{ij}$, Value = LOSER \\
	\> \> \> else: \\
      \> \> \> \> for each $V$ in Multivalue: \\
     \> \> \> \> \> output Key = $E_{ij}$, Value = NULL \\
\> 3 Collate: \> Edge as key \\

@@

 \caption{MapReduce algorithm for finding a maximal independent set of
 graph vertices.}

@@

Before the iterations begin, a random value is assigned to each
vertex, which will be used to compare pairs of vertices.  The vertex
ID itself could be used as the random value, but since the vertices
eventually selected for the MIS depend on the randomization, this may
introduce an unwanted bias.  Instead a random number generator can be
used to assign a random value $R_i$ to each vertex (during a map()
operation), which is then carried along with the vertex ID through
each stage of the algorithm.  In the notation of Figure
\ref{fig:luby}, each $V_i$ is then really two quantities, a vertex ID
(1 to $N$) and $R_i$.  Alternatively, the $R_i$ can be used to relabel
the vertices in a random manner, assigning each a new, unique vertex
ID from 1 to $N$.  In this case, $V_i$ is simply the new vertex ID,
and the vertices in the final MIS could be remapped to recover their
original IDs.  These operations can be performed with a few extra
MapReduce stages, outside the iteration loop.

The pre-iteration clone() operation converts the initial list of edge
key/values one by one to key/multivalue pairs, with the edge as the
key, and NULL as the value.  This is so that the iteration can begin
with a reduce() operation, without need for a collate() operation.  A
second empty MapReduce object is also initialized, which will
accumulate MIS vertices as the iterations proceed.

At each iteration, the current set of edges are examined to identify
winning vertices.  There are removed and added to the MIS.  Vertices
adjacent to winners are flagged as losers, and all edges adjacent to
losers are flagged for removal at the start of the next iteration.
Thus each iteration removes edges from the initially full graph, and
adds vertices to the MIS.  When the graph is empty, the MIS is
complete.

The first reduce() operation flags the two vertices in each edge as a
potential winner or loser.  This is done by comparing the two random
values for the vertices (or the randomized vertex IDs as explained
above).  The first collate() operation thus produces a multivalue for
each vertex which contains all the vertices it shares an edge with and
associated winner/loser flags for $V_i$.  In the second reduce()
operation, if $V_i$ won the comparison with all its edge neighbors,
then it is a ``winner'' vertex and is emitted to the second MapReduce
object.  All vertices connected to the winner are emitted with a loser
flag.  If $V_i$ is not an overall winner, its edge vertices are
emitted without the loser flag.  The second collate() operation
collects this new information for each vertex.

In the third reduce() operation, if $V_i$ was flagged as a loser by
any edge neighbor who was a winner, then it becomes a loser.  All of
its edges are emitted with a loser flag.  If $V_i$ was not flagged as
a loser, its edges are emitted with no flag.  The final collate()
operation gathers this information for each edge; each edge will have
two values (from each vertex), which are either a loser flag or NULL.
At the first stage of the next iteration, any edge flagged as a loser
does not compare its two vertices; it is effectively deleted from the
graph.

The Luby iterations end when all edges have been removed from the
graph.  At this point, the second MapReduce object contains a complete
MIS of vertices.

\subsection{Single source shortest path}
